---
title: "Binary Anchored STC Analysis Example"
author: "Unanchored STC Analysis Package"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: bootstrap
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.width = 10, fig.height = 6)
```

# Binary Anchored STC Analysis Example

This example demonstrates how to perform binary anchored STC analysis using the functions in `binary_astc_analysis.R` and `binary_astc_html_reporting.R`.

Based on NICE DSU Technical Support Document 18 methodology.

## Setup and Data Preparation

```{r load-packages}
# Load required packages and functions
library(dplyr)
library(tidyr)

# Source the analysis and reporting functions
source("binary_astc_analysis.R")
source("binary_astc_html_reporting.R")

# Set seed for reproducibility
set.seed(42)
```

## Example Data Generation

```{r generate-ipd-data}
cat("Generating example data for anchored STC analysis...\n")

# Example 1: IPD from AB trial (Treatment A vs Treatment B)
# This represents individual patient data we have access to
n_ab <- 400
ipd_data <- data.frame(
  patient_id = 1:n_ab,
  treatment = rep(c("A", "B"), each = n_ab/2),
  age = c(rnorm(n_ab/2, mean = 65, sd = 10), rnorm(n_ab/2, mean = 63, sd = 10)),
  gender = c(rbinom(n_ab/2, 1, 0.6), rbinom(n_ab/2, 1, 0.65)),
  smoking = c(rbinom(n_ab/2, 1, 0.3), rbinom(n_ab/2, 1, 0.35))
)

# Generate outcome based on covariates and treatment
ipd_data$outcome_prob <- with(ipd_data, {
  linear_pred <- -1.2 + 
    0.02 * (age - 65) +           # Age effect (centered at 65)
    0.5 * gender +                # Gender effect (male = 1)
    0.3 * smoking +               # Smoking effect
    ifelse(treatment == "B", -0.8, 0)  # Treatment B effect
  
  # Add some treatment interactions
  linear_pred <- linear_pred + 
    ifelse(treatment == "B", 0.015 * (age - 65), 0)  # Age-treatment interaction
  
  1 / (1 + exp(-linear_pred))
})

ipd_data$outcome <- rbinom(n_ab, 1, ipd_data$outcome_prob)
ipd_data$outcome_prob <- NULL  # Remove probability column

# Convert gender and smoking to more intuitive coding
ipd_data$male <- ipd_data$gender
ipd_data$smoker <- ipd_data$smoking
ipd_data <- ipd_data[, c("patient_id", "treatment", "age", "male", "smoker", "outcome")]
```

### IPD Data Summary

```{r summarize-ipd}
cat("IPD Data Summary:\n")
print(ipd_data %>% 
  group_by(treatment) %>% 
  summarise(
    n = n(),
    age_mean = mean(age),
    male_prop = mean(male),
    smoker_prop = mean(smoker),
    outcome_events = sum(outcome),
    outcome_rate = mean(outcome),
    .groups = "drop"
  ))
```

## Example Aggregate Data

```{r define-aggregate-data}
# Example 2: Aggregate data from AC trial (Treatment A vs Treatment C)
# This represents published summary statistics from another trial

# Example aggregate data summaries
agd_covariate_summaries <- list(
  age = list(mean = 68.5, sd = 12.2),      # Older population
  male = list(mean = 0.45, sd = NULL),     # Lower proportion of males
  smoker = list(mean = 0.25, sd = NULL)    # Lower smoking rate
)

# Outcome data for AC trial
agd_outcome_summaries <- list(
  A = list(events = 45, total = 150),      # Reference arm A
  C = list(events = 38, total = 150)       # New treatment C
)

# Create covariate mapping (maps IPD variables to AgD summaries)
covariate_mapping <- list(
  age = "age",
  male = "male", 
  smoker = "smoker"
)

# Define multiple models with different covariate adjustments
models <- list(
  "Naive" = character(0),                    # No covariates
  "Age_Adjusted" = c("age"),                 # Age only
  "Demographics" = c("age", "male"),         # Age + gender
  "Full_Model" = c("age", "male", "smoker")  # All covariates
)

cat("\nAggregate Data (AC Trial) Summary:\n")
cat("Treatment A: 45/150 events (30.0%)\n")
cat("Treatment C: 38/150 events (25.3%)\n")
cat("Mean age: 68.5 years\n")
cat("Male proportion: 45%\n")
cat("Smoking proportion: 25%\n")

cat("\nModels to be fitted:\n")
for (i in seq_along(models)) {
  model_name <- names(models)[i]
  covariates <- models[[i]]
  if (length(covariates) == 0) {
    cat("  ", model_name, ": No covariates (naive model)\n")
  } else {
    cat("  ", model_name, ": ", paste(covariates, collapse = ", "), "\n")
  }
}
```

## Anchored STC Analysis

```{r run-anchored-stc}
# Run analysis with bootstrap (enhanced for debugging)
cat("\n================================================================================ \n")
cat("PERFORMING ANCHORED STC ANALYSIS\n")
cat("================================================================================ \n")

results <- analyze_anchored_stc(
  ipd_data = ipd_data,
  agd_covariate_summaries = agd_covariate_summaries,
  agd_outcome_summaries = agd_outcome_summaries,
  outcome_col = "outcome",
  treatment_col = "treatment",
  covariate_mapping = covariate_mapping,
  models = models,
  reference_arm = "A",
  ipd_comparator_arm = "B",
  agd_comparator_arm = "C",
  use_bootstrap = TRUE,  # Enable bootstrap
  n_bootstrap = 20,      # Reduced for debugging
  include_interactions = TRUE
)
```

### Bootstrap Debugging

```{r bootstrap-debug}
# DEBUG: Check bootstrap results for first model
cat("\n================================================================================ \n")
cat("BOOTSTRAP DEBUG INFO\n")
cat("================================================================================ \n")

first_model <- names(models)[1]
bootstrap_results <- results$model_results[[first_model]]$bootstrap_cis

if (!is.null(bootstrap_results)) {
  cat("Bootstrap results found for model:", first_model, "\n")
  cat("Indirect OR CI:", bootstrap_results$indirect_effect$or_ci, "\n")
  cat("Indirect RR CI:", bootstrap_results$indirect_effect$rr_ci, "\n")
  cat("IPD OR CI:", bootstrap_results$ipd_effect$or_ci, "\n")
  cat("Successful samples:", bootstrap_results$n_successful, "out of", bootstrap_results$n_bootstrap, "\n")
} else {
  cat("ERROR: No bootstrap results found!\n")
}

# Display results
print(results)
```

## Results Summary

```{r results-summary}
cat("\n", paste(rep("=", 80), collapse = ""), "\n")
cat("ANALYSIS RESULTS SUMMARY\n")
cat(paste(rep("=", 80), collapse = ""), "\n")

# Extract key results from each model
model_names <- names(results$model_results)

cat("\n1. Model Comparison Summary:\n")
cat("   Model Name        | Type     | Covariates           | Indirect OR (95% CI)\n")
cat("   ", paste(rep("-", 70), collapse = ""), "\n")

for (model_name in model_names) {
  model_data <- results$model_results[[model_name]]
  indirect <- model_data$indirect_comparison
  
  # Format covariates
  if (length(model_data$covariates) == 0) {
    covariates_str <- "None"
  } else {
    covariates_str <- paste(model_data$covariates, collapse = ", ")
  }
  
  # Format type
  type_str <- ifelse(model_data$is_naive, "Naive", "Adjusted")
  
  cat("   ", sprintf("%-16s", model_name), "| ", sprintf("%-8s", type_str), 
      "| ", sprintf("%-20s", covariates_str), 
      "| ", sprintf("%.3f (%.3f, %.3f)", indirect$or, indirect$or_ci_lower, indirect$or_ci_upper), "\n")
}
```

### Detailed Results

```{r detailed-results}
cat("\n2. Aggregate Data Trial Result:\n")
agd_effect <- results$agd_effect
cat("   ", results$study_characteristics$agd_trial, "\n")
cat("   OR (95% CI):", sprintf("%.3f (%.3f, %.3f)", agd_effect$or, agd_effect$or_ci_lower, agd_effect$or_ci_upper), "\n")

cat("\n3. IPD Trial Results (", results$study_characteristics$ipd_trial, "):\n")
# Show results from the full model
full_model_data <- results$model_results[["Full_Model"]]
if (!is.null(full_model_data$enhanced_effect)) {
  effect <- full_model_data$enhanced_effect
  cat("   RR (95% CI):", effect$rr_formatted, "\n")
  cat("   OR (95% CI):", effect$or_formatted, "\n")
  cat("   RD (95% CI):", effect$rd_formatted, "\n")
  cat("   NNT (95% CI):", effect$nnt_formatted, "\n")
}
```

### Clinical Interpretation

```{r clinical-interpretation}
cat("\n4. Clinical Interpretation:\n")

# Use primary indirect comparison for interpretation
primary_indirect <- results$indirect_comparison

interpretation <- if (primary_indirect$or > 1.2) {
  "Treatment C appears to have higher odds of the outcome compared to Treatment B"
} else if (primary_indirect$or < 0.8) {
  "Treatment C appears to have lower odds of the outcome compared to Treatment B"  
} else {
  "No substantial difference detected between Treatment C and Treatment B"
}

significance <- if (primary_indirect$p_value < 0.05) {
  "statistically significant"
} else {
  "not statistically significant"
}

cat("   Primary result is", significance, "(p =", sprintf("%.4f", primary_indirect$p_value), ")\n")
cat("  ", interpretation, "\n")

# Model comparison
cat("\n5. Model Comparison:\n")
naive_result <- results$model_results[["Naive"]]$indirect_comparison
full_result <- results$model_results[["Full_Model"]]$indirect_comparison

or_difference <- abs(full_result$or - naive_result$or)
cat("   Naive OR:", sprintf("%.3f", naive_result$or), "\n")
cat("   Fully adjusted OR:", sprintf("%.3f", full_result$or), "\n")
cat("   Absolute difference:", sprintf("%.3f", or_difference), "\n")

if (or_difference > 0.2) {
  cat("   → Substantial difference detected - covariate adjustment is important\n")
} else {
  cat("   → Minimal difference detected - populations may be similar\n")
}
```

## HTML Report Generation

```{r html-report}
cat("\n", paste(rep("=", 80), collapse = ""), "\n")
cat("GENERATING HTML REPORT\n")
cat(paste(rep("=", 80), collapse = ""), "\n")

# Generate comprehensive HTML report
report_path <- generate_binary_astc_html_report(
  results = results,
  project_name = "Example_Anchored_STC",
  output_dir = "reports",
  title = "Binary Anchored STC Analysis - Multiple Models Example"
)

cat("\nHTML report generated:", report_path, "\n")
```

## Model Validation

```{r model-validation}
cat("\n", paste(rep("=", 80), collapse = ""), "\n")
cat("MODEL VALIDATION\n")
cat(paste(rep("=", 80), collapse = ""), "\n")

# Check model assumptions and fit for each model
for (model_name in model_names) {
  model_data <- results$model_results[[model_name]]
  
  cat("\n", model_name, "Model:\n")
  
  # Model convergence
  if (model_data$model$converged) {
    cat("   ✓ Model converged successfully\n")
  } else {
    cat("   ✗ Model failed to converge\n")
  }
  
  # Covariates
  if (length(model_data$covariates) == 0) {
    cat("   Covariates: None (naive model)\n")
  } else {
    cat("   Covariates:", paste(model_data$covariates, collapse = ", "), "\n")
  }
  
  # Model fit statistics
  reg_summary <- summary(model_data$model)
  cat("   AIC:", sprintf("%.2f", reg_summary$aic), "\n")
  cat("   Deviance:", sprintf("%.2f", reg_summary$deviance), "\n")
}

cat("\n2. Sample Sizes:\n")
cat("   IPD Trial (AB):", nrow(ipd_data), "patients\n")
cat("   AgD Trial (AC):", sum(sapply(agd_outcome_summaries, function(x) x$total)), "patients\n")

cat("\n3. Event Rates:\n")
ipd_summary <- results$data_summaries$ipd_summary
for (i in 1:nrow(ipd_summary)) {
  arm <- ipd_summary[i, ]
  cat("   IPD -", arm[[1]], ":", sprintf("%.1f%%", arm[[4]] * 100), 
      sprintf("(%d/%d)", arm[[3]], arm[[2]]), "\n")
}

agd_summary <- results$data_summaries$agd_summary  
for (i in 1:nrow(agd_summary)) {
  arm <- agd_summary[i, ]
  cat("   AgD -", arm[[1]], ":", sprintf("%.1f%%", arm[[4]] * 100), 
      sprintf("(%d/%d)", arm[[2]], arm[[3]]), "\n")
}

cat("\n4. Covariate Centering:\n")
# Check centering from first adjusted model
full_model <- results$model_results[["Full_Model"]]$model
if (!is.null(full_model$covariate_centers)) {
  for (cov in names(full_model$covariate_centers)) {
    center_val <- full_model$covariate_centers[[cov]]
    cat("  ", cov, "centered at:", sprintf("%.3f", center_val), "\n")
  }
} else {
  cat("   No covariates were centered\n")
}
```

## Analysis Notes

### Anchored STC Method

- Uses individual patient data from AB trial and aggregate data from AC trial
- Connects trials through shared reference arm A
- Adjusts for population differences by centering IPD covariates on AgD means
- Accounts for effect modification through treatment-covariate interactions
- Provides indirect comparison of C vs B adjusted for population differences

### Key Assumptions

- Shared reference arm A has similar characteristics across trials
- Effect modifiers are correctly identified and measured
- No unmeasured confounders that differ between populations
- Logistic regression model is correctly specified
- Treatment-covariate interactions are appropriately modeled

### Interpretation Guidelines

- Results represent the treatment effect in the aggregate data population
- Confidence intervals account for uncertainty in both trials
- P-values test the null hypothesis of no difference between C and B
- Consider clinical significance in addition to statistical significance

```{r completion-message}
cat("\n", paste(rep("=", 80), collapse = ""), "\n")
cat("ANALYSIS COMPLETED SUCCESSFULLY\n")
cat("Report saved to:", report_path, "\n")
cat(paste(rep("=", 80), collapse = ""), "\n")
```

## Session Information

```{r session-info}
sessionInfo()
``` 